Comparative effectiveness of BNT162b2 versus mRNA-1273 boosting in England: a cohort study in OpenSAFELY-TPP

### 2022-07-06 write-up/manuscript.Rmd
rmds

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmitted",
      #"covidadmittedproxy1",
      #"covidcritcare",
      "coviddeath"
      #"noncoviddeath"
    ) %>% 
    set_names(., .)

@@

matchset <- "B"

@@

*Results* `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(n))` people were matched in each group, contributing a total of `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up. The 12-week risk per `r perN_format` people of positive SARS-CoV-2 test for those receiving BNT162b2 and mRNA-1273 was `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` (RD `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)`; HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)`) respectively.

@@

The UK COVID-19 vaccination programme delivered its first "booster" doses in September 2021 @NHSEnglandNHS. Following guidance from the Joint Committee for Vaccination and Immunisation (JCVI) expert advisory group@covid19 boosting followed a staged roll-out that prioritised groups at high risk of infection or severe disease, and was then progressively extended to the whole adult population by mid December 2021 @NHSEnglandNHSb @NHSEnglandNHSa1 @NHSEnglandNHSa2. The BNT162b2 Pfizer-BioNTech vaccine was used initially for boosting, with a half-dose of Moderna mRNA-1273 subsequently also used. The concurrent administration of the BNT162b2 and mRNA-1273 vaccines for boosting in England, receipt of which was largely determined by local availability rather than any clinical criteria, enables a direct comparison of their effectiveness. No randomised trials have made such a comparison for efficacy outcomes.

On behalf of NHS England, we used the OpenSAFELY-TPP database, covering 40% of English primary care practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to compare the effectiveness of BNT162b2 and mRNA-1273 boosting in adult recipients.

@@

We considered all adults (aged 18 years and over) who received a booster dose of BNT162b2 or mRNA-1273 between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyend_date), "%e %B %Y")` inclusive, during which time both vaccine brands were being used. People were eligible if they: were registered at a GP practice using TPP's SystmOne clinical information system at the time of boosting; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and mRNA-1273 were not considered due to small numbers); were not a health or social care worker, not resident in a care or nursing home and not medically housebound or receiving end-of-life care; had no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; were not hospitalised at the time of boosting; and had complete information on sex, deprivation, and NHS region (defined using Sustainability and Transformation Partnerships (STP, a geographical grouping of NHS and local authorities).

@@

BNT162b2 booster recipients were matched with mRNA-1273 booster recipients in a one to one ratio without replacement, on the following characteristics: date of booster dose; primary vaccine course (BNT162b2 or ChadOx1-S); Date of second vaccine dose (7 day caliper); sex (male or female); age (3 year caliper and within age groups defined by JVCI risk groups); Clinical risk group defined by JCVI (clinically extremely vulnerable, clinically at-risk, neither); Index of Multiple Deprivation (IMD, grouped by quintile); STP as a surrogate for geographical region; evidence of prior SARS-CoV-2 infection (any of positive SARS-CoV-2 test, probable infection documented in primary care, or COVID-19 hospital attendance or admission); morbidity count (0, 1, or 2 or more of diabetes, BMI over 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease). The supplementary materials provide more information on how these characteristics were defined.

<!-- #### Matching strategies -->

<!-- ```{r, table-jcvi, echo=FALSE, message=FALSE, warning=FALSE} -->

<!-- tribble( -->
<!--   ~Factor, ~Description, ~A, ~B, -->

<!--   "Date of booster dose",  "Matched exactly",   -->
<!--     "X", "X", -->
<!--   "Primary course vaccine brand", "BNT162b2; ChadOx1-S",  -->
<!--     "X", "X", -->
<!--   "Date of second vaccine dose", "7 day caliper",  -->
<!--     "X", "X", -->
<!--   "Age",  "3 year caliper",  -->
<!--     "X", "X", -->
<!--   "JCVI age group", "18-39; 40-49; 50-54; 55-59; 60-64; 65-69; 70-75; 75-79; 80 and over", -->
<!--     "X", "X", -->
<!--   "JCVI Clinical risk group", "clinically extremely vulnerable; clinically at-risk; neither", -->
<!--     "X", "X", -->
<!--   "English Index of Multiple Deprivation", "Grouped by quintile", -->
<!--     "X", "X", -->
<!--   "NHS region", "East of England; Midlands; London; North East and Yorkshire; North West; South East; South West", -->
<!--     "X", "", -->
<!--   "Sustainability and Transformation Partnership (STP)", "a geographical grouping of NHS and local authorities", -->
<!--     "", "X", -->
<!--   "Sex", "Male; Female", -->
<!--     "", "X",  -->
<!--   "Evidence of prior SARS-CoV-2 infection", 'Any of: positive SARS-CoV-2 test; "probable" infection documented in primary care; COVID-19 hospital admission', -->
<!--     "", "X", -->
<!--   "Morbidity count", "0, 1, or 2 or more of the following conditions: diabetes; BMI over 40kg/m^2^; chronic heart disease; chronic kidney disease; chronic liver disease; chronic respiratory disease or severe asthma; chronic neurological disease",  -->
<!--     "", "X" -->
<!-- ) %>% -->
<!-- gt() %>% -->
<!-- tab_spanner( -->
<!--   label = "Matching strategy", -->
<!--   columns = all_of(c("A", "B")) -->
<!-- ) %>% -->
<!-- fmt_markdown(columns = TRUE) -->

<!-- cat("  \n\n") -->
<!-- ``` -->

@@

We considered three effectiveness outcomes. Positive SARS-CoV-2 test, identified using SGSS testing records and based on swab date. Both polymerase chain reaction (PCR) and lateral flow test results are included, without differentiation between symptomatic and asymptomatic infection. COVID-19 hospital admission, identified using HES in-patient hospital records with U07.1 or U07.2 ICD-10 reason for admission codes @emergenc. COVID-19 death, defined as deaths with U07.1 or U07.2 ICD-10 codes mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).

@@

Each person was followed from the day of boosting (i.e., time zero) until the earliest of the outcome of interest, death, practice de-registration, 12 weeks, or `r format(as.Date(study_dates$followupend_date), "%e %B %Y")`.

@@

We estimated the cumulative incidence in each treatment group using the Aalen-Johansen estimator. COVID-19 and non-COVID-19 deaths were treated as competing events. Risk differences (RD) and hazard ratios (RRs) comparing the cumulative incidences were derived at 12 weeks. We also estimated hazard ratios (HRs) comparing booster vaccine groups using Cox models.

@@

We estimated comparative effectiveness separately in the following subgroups: primary vaccine course (ChAdOx1-S or BNT162b2); age (18-64 or $\ge$ 65 years); clinical vulnerability, as defined by JCVI (Not clinical at-risk, clinically at-risk, or clinically extremely vulnerable); evidence of prior SARS-CoV-2 infection or not.

@@

To satisfy strict re-identification minimisation requirements for statistical outputs from OpenSAFELY's secure environment, all counts were reported to the nearest $6n-3$ ($3, 9, 15, 21, ...$). Zero counts were unchanged. Similarly, the Aalen-Johansen cumulative incidence estimates were rounded such that each "step" is based on at least 6 events.

@@

`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(n))` were matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7B", vax3_type=="moderna") %>% pull(pct_step))` respectively (Figure 1).

@@

There were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(n.event)` positive SARS-CoV-2 tests, `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospitalisations, and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths at 12-weeks, across `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up (Table 3).

Absolute risks at 12 weeks of positive SARS-CoV-2 test for those receiving BNT162b2 and mRNA-1273 were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people respectively, representing a risk difference of `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` and a hazard ratio of `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)` (Table 3, Figure 3). Corresponding 12-week risks for COVID-19 hospitalisation were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(coxhrCI)`), and for COVID-19 death were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(coxhrCI)`).

@@

A notable exception was outcomes stratified by primary vaccination series: the estimated benefit of mRA-1273 over BNT162b2 was higher amongst those who received BNT162b2 for their first two vaccine doses (HR for positive SARS-CoV-2 `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="vax12_type", subgroup_level=="pfizer-pfizer") %>% pull(coxhrCI)`) compared with ChAdOx1 (HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(coxhrCI)`).

@@

This observational study in over 4 million adults compares the effectiveness of BNT162b2 and mRNA-1273 vaccines against COVID19-related outcomes in the national booster vaccination programme in England. We estimated that mRNA-1273 provided slightly more protection than BNT162b2 did against a positive SARS-CoV-2 test (RD = `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` per `r perN_format`) and hospitalisation (RD = `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI_95)`) in the first 12 weeks after boosting. For COVID-19 death, the difference was equivocal (RD = `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI_95)`) though there were only `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` events across both vaccine types, limiting statistical power.

@@

```{r table-table1, echo=FALSE, warning=FALSE, message=FALSE}
table1A <- 
  matchA_table1 %>%
  filter(!is.na(`stat_1`)) %>%
  group_by(variable) %>%
  mutate(
    nlabel=n(),
    #label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
    var_label = if_else(row_number()==1, var_label, ""),
    label = if_else(nlabel!=1, label, ""),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(matchA_table1by$by_col, matchA_table1by$by_chr))

table1B <- 
  matchB_table1 %>%
  filter(!is.na(`stat_1`)) %>%
  group_by(variable) %>%
  mutate(
    nlabel=n(),
    #label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
    var_label = if_else(row_number()==1, var_label, ""),
    label = if_else(nlabel!=1, label, ""),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(matchB_table1by$by_col, matchB_table1by$by_chr))

@@

table1B %>%
  gt() %>%
  fmt_markdown(columns = "label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
    var_label = "Variable",
    #`BNT162b2_A` = "BNT162b2",
    #`Moderna mRNA-1273_A` = "mRNA-1273",
    `BNT162b2` = "BNT162b2",
    `Moderna mRNA-1273` = "mRNA-1273",
    label=md("Characteristic")
  ) %>%
  # tab_spanner(
  #   label = "Matching strategy A",
  #   columns = ends_with("_A")
  # ) %>%
  # tab_spanner(
  #   label = "Matching strategy B",
  #   columns = ends_with("_B")
  # ) %>%
  cols_align(
    align = c("right"),
    columns = (c("BNT162b2", "Moderna mRNA-1273"))
  ) %>%
  cols_hide(c("variable", "order")) %>%
  tab_options(
    table.font.size = 8
  )

@@

contrasts_overall_table %>%
  filter(
    outcome %in% outcomes,
    strategy=="B"
  ) %>%
  group_by(outcome_descr) %>%
  select(
    outcome_descr,
    subgroup_descr,
    subgroup_level_descr,

@@

data_annotate <- tibble(
    vax3_date = date(c("2021-11-26", "2021-11-26", "2022-01-07", "2022-01-07")),
    cumuln = c(1000000, -2900000, 0, -2900000),
    label = c("mRNA-1273", "BNT162b2", "Matched", "Unmatched"),
    colour = c(brewer_pal(type="qual", palette="Set2")(2)[c(2,1)], "darkgrey", "grey"),
    vjust = c(1.2,-0.1,1.2,-0.1)
  )

>>>>

  outcomes <- 
    c(
      "postest",
      "covidemergency",
      "covidadmitted",
      #"covidadmittedproxy1",
      #"covidcritcare",
      "coviddeath"
      #"noncoviddeath"
    ) %>% 
    set_names(., .)

@@

matchset <- "A"

@@

*Results* `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(n))` people were matched in each group, contributing a total of `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up. The 12-week risk per `r perN_format` people of positive SARS-CoV-2 test for those receiving BNT162b2 and mRNA-1273 was `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` (RD `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)`; HR `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)`) respectively.

@@

The UK COVID-19 vaccination programme delivered its first "booster" doses in September 2021, initially in groups at high risk of severe disease then across the adult population. A half-dose of the BNT162b2 Pfizer-BioNTech vaccine was used initially, with Moderna mRNA-1273 subsequently also used. The concurrent administration of the BNT162b2 and mRNA-1273 vaccines for boosting in England, receipt of which was largely determined by local availability rather than any clinical criteria, enables a direct comparison of their effectiveness. No randomised trials have made such a comparison.

We used the OpenSAFELY-TPP database, covering 40% of English primary care practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to compare the effectiveness of BNT162b2 versus mRNA-1273 in boosted adults between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyend_date), "%e %B %Y")`, during which time both vaccine brands were being used.

@@

We considered all adults aged $\ge$ 18 years who received a booster dose of BNT162b2 or mRNA-1273 between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyend_date), "%e %B %Y")` inclusive (the "recruitment period"). People were eligible if they: were registered at a GP practice using TPP's SystmOne clinical information system at the time of boosting; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and mRNA-1273 were not considered due to small numbers); were not a health or social care worker, not resident in a care or nursing home and not medically housebound or receiving end-of-life care; had no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; were not undergoing an unplanned hospital admission; and had complete information on sex, ethnicity, deprivation, and NHS region.

@@

BNT162b2 booster recipients were matched 1:1 with mRNA-1273 booster recipients without replacement. Two different matching strategies were used: one simpler strategy resulting in more matches but with greater imbalance between booster vaccine groups, and another more complex strategy resulting in fewer matches but with less imbalance.

#### Matching strategies

```{r, table-jcvi, echo=FALSE, message=FALSE, warning=FALSE}

tribble(
  ~Factor, ~Description, ~A, ~B,

  "Date of booster dose",  "Matched exactly",  
    "X", "X",
  "Primary course vaccine brand", "BNT162b2; ChadOx1-S", 
    "X", "X",
  "Date of second vaccine dose", "7 day caliper", 
    "X", "X",
  "Age",  "3 year caliper", 
    "X", "X",
  "JCVI age group", "18-39; 40-49; 50-54; 55-59; 60-64; 65-69; 70-75; 75-79; 80 and over",
    "X", "X",
  "JCVI Clinical risk group", "clinically extremely vulnerable; clinically at-risk; neither",
    "X", "X",
  "English Index of Multiple Deprivation", "Grouped by quintile",
    "X", "X",
  "NHS region", "East of England; Midlands; London; North East and Yorkshire; North West; South East; South West",
    "X", "",
  "Sustainability and Transformation Partnership (STP)", "a geographical grouping of NHS and local authorities",
    "", "X",
  "Sex", "Male; Female",
    "", "X", 
  "Evidence of prior SARS-CoV-2 infection", 'Any of: positive SARS-CoV-2 test; "probable" infection documented in primary care; COVID-19 hospital admission',
    "", "X",
  "Morbidity count", "0, 1, or 2 or more of the following conditions: diabetes; BMI over 40kg/m^2^; chronic heart disease; chronic kidney disease; chronic liver disease; chronic respiratory disease or severe asthma; chronic neurological disease", 
    "", "X"
) %>%
gt() %>%
tab_spanner(
  label = "Matching strategy",
  columns = all_of(c("A", "B"))
) %>%
fmt_markdown(columns = TRUE)

Morbidity count included the following conditions: diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease"

@@

We considered four effectiveness outcomes. Positive SARS-CoV-2 test, identified using SGSS testing records and based on swab date. Both polymerase chain reaction (PCR) and lateral flow test results are included, without differentiation between symptomatic and asymptomatic infection. COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc. COVID-19 hospital admission, identified using HES in-patient hospital records with U07.1 or U07.2 ICD-10 reason for admission codes. COVID-19 death. Deaths were classified as from COVID-19 if deaths with the U07.1 or U07.2 ICD-10 codes were mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).

@@

Each person was followed from the day of boosting (i.e., time zero) until the earliest of outcome, death, practice de-registration, 12 weeks, or `r format(as.Date(study_dates$followupend_date), "%e %B %Y")`.

@@

We estimated the cumulative incidence in each treatment group using the Aalen-Johansen estimator, with COVID-19 and non-COVID-19 treated as competing events (when these were not the event of interest). Risk differences (RD) and hazard ratios (RRs) comparing the cumulative incidences were derived at 12 weeks. We also estimated hazard ratios (HRs) comparing booster vaccine groups using Cox models.

@@

We estimated comparative effectiveness separately in the following subgroups: primary vaccine course (ChAdOx1-S or BNT162b2); age (18-64 or $\ge$ 65 years); clinical vulnerability, as defined by the UK Joint Committee for Vaccination and immunisation (JCVI) (Not clinical at-risk, clinically at-risk, or clinically extremely vulnerable); evidence of prior SARS-CoV-2 infection or not.

@@

To reduce re-identification risk, counts are reported to the nearest $6n-3$ ($3, 9, 15, 21, ...$). Event times were deferred until there were at least 6 events, so that each "step" in the Aalen-Johansen cumulative incidence estimates is based on at least 6 events. An offset of -3 is then applied to de-bias this rounding.

@@

For matching strategy A `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7A", vax3_type=="pfizer") %>% pull(n))` were matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7A", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7A", vax3_type=="moderna") %>% pull(pct_step))` respectively. For matching strategy B `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(n))` were matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7B", vax3_type=="moderna") %>% pull(pct_step))` respectively (Figure 1).

@@

For matching strategy A, there were `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(n.event)` positive SARS-CoV-2 tests, `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospitalisations, and `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths at 12-weeks, across `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up (Table 3).

Absolute risks at 12 weeks of positive SARS-CoV-2 test for those receiving BNT162b2 and mRNA-1273 were `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people respectively, representing a risk difference of `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` and a hazard ratio of `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)` (Table 3, Figure 3). Corresponding 12-week risks for COVID-19 hospitalisation were `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(coxhrCI)`), and for COVID-19 death were `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(coxhrCI)`).

For severe outcomes, RDs and HRs were similar for matching strategy B where treatment groups were more closely matched. The risk of positive SARS-CoV-2 test was more similar between BNT162b2 and mRNA-1273 when estimated under matching strategy B, with `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` events per `r perN_format` people respectively, (RD `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)`; HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)`).

@@

A notable exception was outcomes stratified by primary vaccination series: the estimated benefit of mRA-1273 over BNT162b2 was higher amongst those who received BNT162b2 for their first two vaccine doses (HR for positive SARS-CoV-2 `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="vax12_type", subgroup_level=="pfizer-pfizer") %>% pull(coxhrCI)`) compared with ChAdOx1 (HR `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(coxhrCI)`).

@@

This observational study in over 4 million adults compares the effectiveness of BNT162b2 and mRNA-1273 vaccines against COVID19-related outcomes in the national booster vaccination programme in England. We estimated that mRNA-1273 provided slightly more protection than BNT162b2 did against a positive SARS-CoV-2 test (RD = `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` per `r perN_format`) and hospitalisation (RD = `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI_95)`) in the first 12 weeks after boosting. For COVID-19 death, the difference was equivocal (RD = `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI_95)`) though there were only `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` events across both vaccine types, limiting statistical power.

@@

```{r table-table1, echo=FALSE, warning=FALSE, message=FALSE}
table1A <- 
  matchA_table1 %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(matchA_table1by$by_col, matchA_table1by$by_chr))

table1B <- 
  matchB_table1 %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(matchB_table1by$by_col, matchB_table1by$by_chr))

@@

table1 %>%
  gt() %>%
  fmt_markdown(columns = "label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
    `BNT162b2_A` = "BNT162b2",
    `Moderna mRNA-1273_A` = "mRNA-1273",
    `BNT162b2_B` = "BNT162b2",
    `Moderna mRNA-1273_B` = "mRNA-1273",
    label=md("Characteristic")
  ) %>%
  tab_spanner(
    label = "Matching strategy A",
    columns = ends_with("_A")
  ) %>%
  tab_spanner(
    label = "Matching strategy B",
    columns = ends_with("_B")
  ) %>%
  cols_align(
    align = c("right"),
    columns = ends_with(c("_A", "_B"))
  ) %>%
  cols_hide(c("variable", "var_label", "order")) %>%
  tab_options(
    table.font.size = 8
  )

@@

contrasts_overall_table %>%
  filter(
    outcome %in% outcomes,
    strategy=="A"
  ) %>%
  group_by(outcome_descr) %>%
  select(
    outcome_descr,
    subgroup_descr,
    subgroup_level_descr,

@@

plot_coverage_cumuln <-
  match_coverage %>%
  mutate(
   # treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
    #status_descr = fct_rev(fct_recoderelevel(as.character(status), recoder$status)),
    status_descr = fct_rev(status_descr),
    cumuln = cumuln*((treatment*2) - 1)
  ) %>%
  ggplot()+
  geom_col(
    aes(
      x=vax3_date+0.5,
      y=cumuln,
      group=paste0(treatment,status_descr),
      fill=treatment_descr,
      alpha=status_descr,
      colour=NULL
    ),
    position=position_stack(reverse=TRUE),
    width=1
  )+
  geom_rect(xmin=xmin, xmax= xmax+1, ymin=-6, ymax=6, fill="grey", colour="transparent")+
  scale_x_date(
    breaks = seq(xmin, xmax+1, by=14),
    #breaks = unique(lubridate::ceiling_date(match_coverage$vax3_date, "1 month")),
    limits = c(xmin-1, NA),
    labels = scales::label_date("%d/%m"),
    expand = expansion(add=1),
    sec.axis = sec_axis(
      trans = ~as.Date(.),
      breaks=as.Date(seq(floor_date(xmin, "month"), ceiling_date(xmax, "month"),by="month")),
      labels = scales::label_date("%B %y")
    )
  )+
  scale_y_continuous(
    labels = ~scales::label_number(accuracy = 1, big.mark=",")(abs(.)),
    expand = expansion(c(0, NA))
  )+
  scale_fill_brewer(type="qual", palette="Set2")+
  scale_colour_brewer(type="qual", palette="Set2")+
  scale_alpha_discrete(range= c(0.8,0.4))+
  labs(
    x="Date",
    y="Cumulative booster vaccines",
    colour=NULL,
    fill=NULL,
    alpha=NULL
  ) +
  theme_minimal()+
  theme(
    axis.line.x.bottom = element_line(),
    axis.text.x.top=element_text(hjust=0),
    strip.text.y.right = element_text(angle = 0),
    axis.ticks.x=element_line(),
    legend.position = "bottom",
    legend.box = "vertical"
  )+
  NULL
